### 2019-10-09 docs/slides/vertical_intro.Rmd

---
title: "Reproducible psychological research and project management in R"
subtitle: "A solution stack"
author: "Matthew Crump"
date: "2018/07/20 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    css: ["slides_css/defaultb.css", "slides_css/metropolisb.css", "slides_css/metropolis-fontsb.css"]
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE, echo=FALSE}
options(htmltools.dir.version = FALSE)
knitr::opts_chunk$set(echo = FALSE,message=FALSE,warning=FALSE, cache = TRUE)
```

class: pink, center, middle, clear

# Everything in its right place - Radiohead

---

# Hodge Podge Management

Using lots of different programs to accomplish different jobs

1. Experiment builders (E-prime, Presentation, PsychoPy, in-house software, etc.)
2. Data Analysis (Excel, SPSS, R, Python, Matlab, etc.)
3. Write up (Word, Latex, R markdown)
4. Presentation (powerpoint, keynote)

Outcome: lots of different files everywhere, lots of work by hand, **can be very difficult to reconstruct or redo specific steps**

---

# Solution Stack Management

From Wikipedia: In computing, a **solution stack or software stack** is a set of software subsystems or components needed to create a complete platform such that no additional software is needed to support applications. Applications are said to "run on" or "run on top of" the resulting platform.

---

# R as a solution stack

1. Program Experiments (exprmntr and JsPsych)
2. Data analysis (e.g., tidyverse)
3. Graphing (ggplot2)
4. Reproducible reports (R markdown)
5. Slide decks (xaringan)
6. APA papers (papaja)
7. Sharing custom functions (R packages)
8. Sharing everything all in one place (R markdown websites, Github, and pkgdown)

---

# R as a control center

Having a full solution stack in R doesn't mean you can't use other things.

R can often be used as a control center to run other code (e.g., python, C++, etc.)

---