### 2019-11-13 inst/rmarkdown/templates/slidy/skeleton/skeleton.Rmd

---
title: "Example slide presentation using slidy"
subtitle: "A place for a subtitle"
author: "Matthew Crump"
date: "10/31/2019"
output:
  slidy_presentation:
    self_contained: true
    theme: yeti
    incremental: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Example slidy presentation {.middle .center}

Check out the documentation on slidy for more info <https://bookdown.org/yihui/rmarkdown/slidy-presentation.html>

- Note that this slide deck has some custom CSS controlling the styling and look and feel. This can be customized further by editing the style.css file.

## A slide with a header {.bigger}

## A slide with a header {.bigger}

And some text in it.

## A slide with a header and bullet points {.bigger}

- bullet 1
- bullet 2
- bullet 3

## Something in the middle{.middle .center}

## Slide with a picture

```{r, echo = FALSE, out.width="50%", fig.align="center"}
plot(pressure$temperature, pressure$pressure)
```

## The end {.middle .center}

- don't forget there are other R markdown options for slide decks
- eg., xaringan, and ioslides, and probably a few others